---
title: "Troubleshoot for Parsing of Riley Asthmaster Data"
author: Daniel P. Hall Riggins, MD, MPH
date: today
format: html
---

Load Dependencies:

```{r}
library(tidyverse)
```

Read in data dictionary that I annotated with [shorthand for data types](https://readr.tidyverse.org/reference/read_delim.html):

```{r}
dd <- readxl::read_excel("data/modified_elite_list_data_dictionary.xlsx")

glimpse(dd)
```

Concatenate the data types into one string:

```{r}
dd$Type |> 
    cat()
```
Remove spaces in the string to create the parsing specification:

```{r}
parse_spec <-
    "i i T T i T i c c c c c i c l l l l l l l l d d l l i l l l i d d l l i l d d d l l d d l l d d l l d d l l d d l i l l d d l l l l l l d d i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d d i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i d d d d d d d d d d d d d d d i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i i l l l l l i c c c c c c c c c c c c c c c l d d d d d d d d d d d d d d d c c c c c c c c c c c c c c c i i i i i i i i i i i i i i i c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c d d d d d d d d d d d d d d d c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c c d d d d d d d d d d d d d d d c c c c c c c c c c c c c c c" |> 
    str_replace_all("[:space:]", "")
```

Parse the dataset using the specification above:

```{r}
df <- read_csv("data/RileyAsthmaster.csv", col_types = parse_spec)

glimpse(df)
```

Starting at the CBC differential variables (columns 383+), certain variable groups had parsing errors because of limits of detection. To prevent loss of data, I specified these variables as "character" type. 

Using the following pattern, I inspected different variable groups and classified the types of detection limits that were observed:

```{r}
df |> 
    select(FIN, starts_with("Abs_Lymphocytes")) |> 
    pivot_longer(cols = starts_with("Abs_Lymphocytes")) |> 
    select(value) |> 
    arrange(value) |> 
    janitor::tabyl(value) |> 
    select(value, n)
```

Variables with lower limits of detection:

- Abs_Lymphocytes_* variables
- Abs_Neutrophils_* variables
- WBC_* variables

Variables with upper limits of detection:

- Bicarb_* variables
- pCO2_* variables

Variables with both limits of detection:

- Theophylline_Lvl_* variables

Additionally, there were parsing errors with the Lymphocytes_* and Neutrophils_* variables because of certain values reading "see man diff", which below I convert to NA values:

```{r}
parsed_df <-
    df |> 
    mutate(
        across(
            .cols = c(
                starts_with("Lymphocytes"), 
                starts_with("Neutrophils")
            ),
            .fns = ~case_when(
                .x == "see man diff	" ~ NA,
                TRUE ~ as.integer(.x)
            )
        )
    ) 
```

Using this exploratory analysis, I have created the function parse_asthmaster(), which I use to create a raw, parsed dataset in my analysis pipeline.

